---
title: "outlier_exploration"
output: html_document
date: "2025-03-12"
editor_options: 
  markdown: 
    wrap: 72
---

From Ryan

Based on Sabesh's data_clean.rmd code which identified the outliers. The
first few chunks are Sabesh's code. I adjusted the outlier report to
also include IQR, Q1, Q3, and potentially more statistical analyses.
Then its my code.

Gotta import the libraries

```{r import libraries, include=FALSE}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
```

Importing cleaned data as cleaned_df

```{r import dataset}
# Update this variable with the your own filepath
data_filepath <- "/Users/ryandelft/Desktop/MOT143A Business Analytics/Group Project/Github Code/Dataset YT/trending_yt_videos_113_countries_cleaned.csv"

#Import dataset
cleaned_df <- read.csv(data_filepath)
```

```{r basic exploration}
#str(cleaned_df)
summary(cleaned_df)
```

Detecting outliers adapted from Sabesh's code

```{r detect outliers}
detect_outliers <- function(df) {
  numeric_cols <- df %>% 
    select(where(is.numeric))
  outlier_summary <- data.frame(column = character(), lower_bound = numeric(), 
                                upper_bound = numeric(), num_outliers = integer(), 
                                iqr = numeric(), q1 = numeric(), q3 = numeric())
                                
  for (col in colnames(numeric_cols)) {
    q1 <- quantile(numeric_cols[[col]], 0.25, na.rm = TRUE)
    q3 <- quantile(numeric_cols[[col]], 0.75, na.rm = TRUE)
    iqr_value <- q3 - q1
    lower_limit <- q1 - 1.5 * iqr_value
    upper_limit <- q3 + 1.5 * iqr_value
    num_of_outliers <- sum(numeric_cols[[col]] < lower_limit | numeric_cols[[col]] > upper_limit, na.rm = TRUE)
    
    outlier_summary <- rbind(outlier_summary, data.frame(column = col, lower_bound = lower_limit, 
                                                         upper_bound = upper_limit, num_outliers = num_of_outliers, 
                                                         iqr = iqr_value, q1 = q1, q3 = q3))
  }
  return(outlier_summary)
}

```

Printing outlier report

```{r print outlier report}
outlier_report <- detect_outliers(cleaned_df)
print(outlier_report)
```

**Ryan notes**

1\. Shouldn't the upper bounds for daily rank, daily movement, and
weekly movement be 50; lower bounds -49 for movements

2\. Also, like_count, view_count, comment_count lower bounds should also
be 0?

3\. there are also videos which are daily/weekly number 1 with 0 views
and 0 likes - do those need to be deleted?

4\. outliers in daily movement, based on this chart, are videos that
moved up or down by roughly 38 places in a day

**What to do about outliers?**

1\. daily rank: min=0, max=50

2\. daily movement: min=-50, max=50

3\. weekly movemement: min=-50, max=50

4\. view count: min=0, no max

5\. like count: min=0, no max

6\. comment count: min=0, no max

```{r group outliers}
#making df for outliers of a column just to look at them

group_outliers <- function(df) {
  #hard coding the column I want based on outlier metrics
  outlier_rows <- df %>% filter(comment_count >=4041)
   

  return(outlier_rows)
}
```

```{r print outlier_df}
outlier_df <- group_outliers(cleaned_df)
print(outlier_df)
```

#Lets perform a correlation analysis on the outlier_df (From Liam)

```{r correlation outlier_df}
cor_matrix <- outlier_df %>%
  select(view_count, like_count, comment_count, daily_rank, daily_movement, weekly_movement) %>%
  cor(use = "complete.obs")
print(cor_matrix)

```

#Compared to cleaned_df

```{r correlation cleaned_df}
cor_matrix <- cleaned_df %>%
  select(view_count, like_count, comment_count, daily_rank, daily_movement, weekly_movement) %>%
  cor(use = "complete.obs")
print(cor_matrix)

```

Very similar results. But! discrepancy between comment
count/weekly_movement (row/column) in outliers: 0.013 and in cleaned_df:
-0.008. Further, the weekly_movement/comment_count in outliers: 0.0134
and in cleaned_df: -0.0089. There are also discrepancies with the
daily_rank & daily movement wrt comment_count. Comment_count seems like
an interesting predictor of ranking at least.

Interesting relationships between comment_count and daily/weekly
movements when we look at correlation coefficients of the outlier_df and
cleaned_df (full dataset)

Cleaned_df: comment_count / daily_rank: -0.040503626 comment_count /
daily_movement: -0.01030539 comment_count / weekly_movement: -0.00897838

Outlier_df: comment_count / daily_rank: -0.05007134 comment_count /
daily_movement: 0.007202671 comment_count / weekly_movement: 0.01340262
