---
title: "PCA"
output: html_document
date: "2025-03-20"
editor_options: 
  markdown: 
    wrap: 72
---

**We need to do the following:**

1\. Define trendiness (maybe daily_rank \<10 for \>5 days ? something
like that?)

2\. Reduce the dataset with rows that abide to our definition of
trendiness so that we are analyzing actually trendy videos, not all
videos in our dataset

```{r import libraries, include=FALSE}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
```

Import Dataset

```{r import dataset}
# Update this variable with the your own filepath
data_filepath <- "/Users/ryandelft/Desktop/MOT143A Business Analytics/Group Project/Github Code/Dataset YT/trending_yt_videos_113_countries_cleaned.csv"

#Import dataset
dta_draw_from <- read.csv(data_filepath)
```

See sample values

```{r print head}
#head(dta_draw_from)
```

```{r filter trendy videos}
# trendy = daily rank of 10 or higher and a weekly movement of 5 or less
trendy_df_numeric <- dta_draw_from %>%
  filter(daily_rank >= 1 & daily_rank <= 10, weekly_movement < 5)

head(trendy_df_numeric)
```

Change certain non-numeric columns to numeric columns (namely: title,
country, channel_id, publish data, video_id) Will try to do:
description; those will be more complicated, though Therefore excluding:
channel_name (covered with channel_id) Excluding due to null: language,
video_tags

Convert character to numeric

```{r numeric conversion}
#title
trendy_df_numeric$title_numeric <- as.numeric(factor(trendy_df_numeric$title))

#publish date
trendy_df_numeric$date_numeric <- as.numeric(format(ymd_hms(trendy_df_numeric$publish_date), "%Y%m%d"))

#country
trendy_df_numeric$country_numeric <- as.numeric(factor(trendy_df_numeric$country))

#channel_name
trendy_df_numeric$channel_name_numeric <- as.numeric(factor(trendy_df_numeric$channel_name))

#language (not including as there are too many nulls; will delete column)
#trendy_df_numeric$primary_language_numeric <- as.numeric(as.factor(trendy_df_numeric$language))

#video_id
trendy_df_numeric$video_id_numeric <- as.numeric(as.factor(trendy_df_numeric$video_id))
```

Drop character columns already converted

```{r drop character converted}
#trendy_df_numeric <- trendy_df_numeric %>% select(-c(title, publish_date, country, channel_name, language, channel_id, video_id, video_tags, description))
#trendy_df_numeric <- trendy_df_numeric %>% select(-c(primary_language_numeric))
```

Check nulls

```{r check null}
colSums(is.na(trendy_df_numeric))
```

Drop rows with nulls (some in title)

```{r drop null}
trendy_df_numeric <- trendy_df_numeric %>% drop_na(title_numeric)
colSums(is.na(trendy_df_numeric))
```

Standardize

```{r standardize}
trendy_standardized <- scale(trendy_df_numeric[, !names(trendy_df_numeric) %in% "classe" ], center = TRUE, scale = TRUE)
```

Run PCA

```{r pca}
cat("\n====================== Principal Component Analysis of trendy_standardized =====================\n\n")
results <- prcomp(trendy_standardized)
summary(results)
cat("\n================================================================================================\n")

```

Result: PC8 is 90.0%

See which variables are associated with each principle component

```{r pca identification}
cat("\n====================== Principal Component Loadings =====================\n\n")
print(results$rotation)
cat("\n=========================================================================\n")
```

Findings for PC8 loadings: PC8 in order of absolute value: comment_count
0.71, date_numeric 0.38, view_count 0.40, daily_movement 0.27,
weekly_movement 0.17, channel_name_numeric 0.14, title_numeric 0.11 [end
top 8]; daily_rank 0.09, country_numeric 0.08, video_id_numeric 0.05

```{r pca visualization}
library(ggplot2)

visualization_df <- as.data.frame(results$rotation)
visualization_df$variable <- rownames(visualization_df)

ggplot(visualization_df, aes(x = variable, y = PC8)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Feature Importance in PC8")
```

```{r pca contribution again}
library(factoextra)

fviz_pca_var(results, col.var = "contrib")  # Contribution plot
```

From this, we can see the following variables contribute the most (6):
like_count, view_count, comment_count, daily_rank, daily_movement,
weekly_movement The ones contributing the least (5): country_numeric,
title_numeric, channel_name_numeric, date_numeric, video_id_numeric

Drop insignificant columns

```{r drop insignificant}
#dta_draw_from <- dta_draw_from %>% select(-c(country_numeric, title_numeric, channel_name_numeric, date_numeric, video_id_numeric))
#dim(dta_draw_from)
```
