```{r}
# Load required libraries
library(tidyverse)
library(caret)
library(randomForest)

# Load dataset (replace with actual file path)
```


```{r}
youtube_data <- read.csv("C:/Users/keert/OneDrive/Desktop/Business Analytics - Tutorials/trending_yt_videos_113_countries_cleaned.csv")

# View first few rows
head(youtube_data)

# Check for missing values
colSums(is.na(youtube_data))

# Summary statistics
summary(youtube_data)
```

```{r}
set.seed(123)
```

```{r}
colnames(youtube_data)
```
```{r}
# Define engagement score (like-to-view ratio)
youtube_data$engagement_score <- youtube_data$like_count / youtube_data$view_count

# Define trendiness as a function of engagement and movement
youtube_data$trendiness <- ifelse(youtube_data$engagement_score > 0.05 & youtube_data$daily_movement > 0, 1, 0)

# Convert trendiness to a factor for classification
youtube_data$trendiness <- as.factor(youtube_data$trendiness)

# Check distribution
table(youtube_data$trendiness)
```






```{r}
# Convert target variable to factor (classification task)
youtube_data$trendiness <- as.factor(youtube_data$trendiness)

# Select features for modeling
features <- c("view_count", "like_count", "comment_count", 
              "daily_rank", "daily_movement", "weekly_movement")

# Define training dataset
X <- youtube_data[, features]
y <- youtube_data$trendiness
```


```{r}
# Create train-test split 
set.seed(123)
trainIndex <- createDataPartition(y, p = 0.7, list = FALSE)
X_train <- X[trainIndex, ]
X_test <- X[-trainIndex, ]
y_train <- y[trainIndex]
y_test <- y[-trainIndex]
```

```{r}
summary(log_model)
```
`

```{r}
# Logistic Regression Model
#log_model <- train(X_train, y_train, method = "glm", family = "binomial")

# Predict on test data
#log_preds <- predict(log_model, X_test)

# Evaluate model performance
#conf_matrix_log <- confusionMatrix(log_preds, y_test)
#conf_matrix_log

#warning in Logistic regression - Warning: glm.fit: fitted probabilities numerically 0 or 1 occurredWarning: 
#this occurs when there are clear separation between variables.


```{r}
library(randomForest)
```

```{r}
set.seed(123)  # Ensure reproducibility
sample_index <- sample(nrow(youtube_data), size = 0.1 * nrow(youtube_data))  # 10% sample
youtube_sample <- youtube_data[sample_index, ]

# Convert 'trendiness' to a factor for classification
youtube_data$trendiness <- as.factor(youtube_data$trendiness)

# Train Random Forest Model (Classification)
rf_model <- randomForest(trendiness ~ view_count + like_count + comment_count + daily_rank + daily_movement + weekly_movement,
                         data = youtube_data, ntree = 100)

```

```{r}
# Print the Random Forest Model Summary
print(rf_model)
```


```{r}
library(caret)
```

```{r}
# Predict using the test set
rf_preds <- predict(rf_model, X_test)

# Confusion Matrix and Model Evaluation
confusionMatrix(rf_preds, y_test)
```

